---
title: "Data Exploration and Visualization Lab"
output:
  word_document: default
  pdf_document: default
  html_document:
    df_print: paged
---

#### Course Number: IS-470
#### Course Title: Business Intelligence and Big Data
#### Student Name: ADD

# Introduction

The purpose of this lab is to learn the basic of data exploration in R using a few popular packages. The exercise is not meant to be difficult...so take a breath. We are hoping you will have fun and learn something new along the way.
If you get stuck, the examples and questions used within the exercises come from the following articles. Please try it on your own first though and ask help from your classmates.

https://www.pluralsight.com/guides/exploring-data-set-r

https://cran.r-project.org/web/packages/DataExplorer/vignettes/dataexplorer-intro.html

https://r4ds.had.co.nz/exploratory-data-analysis.html

https://rpubs.com/LexyBrooks/Flight-Explore

To complete lab successfully complete the following items listed below:

- Run the code on a data set of your choosing.
- Run the code example code using your data.
- Answer all the questions about the exploratory data analysis you completed on your data.

# Load Packages and Data

The following libraries will be used within the analysis. Remember to install the packages using the `Packages` tab on the right hand side before loading the packages. Also, load in one of the other flight data sets (`airlines`, `airports`, `planes`, `weather`) within this `r` block. 

```{r include=FALSE}
library(tidyr)
library(dplyr)
library(ggplot2)
library(nycflights13)
library(DataExplorer)
library(skimr)

#Using the flights data from the nycflights13 package
flights_data = nycflights13::flights
```

# Data Exploration
Lets take a look at your data and see what we can learn. To get a first glance at my data I decided upon looking at the shape, first few rows, and some summary statistics for the flights data. Used the code that I provided to see what you can find out about your data.

#### Questions

##### How many rows and columns are there in your data set?

**Answer:** 

```{r}
print(dim(flights_data))
```

##### At glance, what types of columns are available in your data set? For example, one observation for the `flights` data is that most of the columns in the data are numeric with a few categorical columns (i.e., origin, dest).

**Answer:** 

```{r}
head(flights_data, 5)
```

##### What additional information did you learn from the summary statistics table? For example, one observation for the `flights` data is that there are few columns that have missing observations and they mostly have to do with features related to flight delay information. 

**Answer:** 

```{r}
summary(flights_data)
```

# Data Visualization
Now that we have a better understanding of the structure of the data and type of data we are working with, let us create a few data visualizations to see if we can learn some more information about our data. First run your data through one of the Exploratory Data Analysis tools loaded: `skimr` or `DataExplorer`. For the example I used `skimr`; however, please feel free to `DataExplorer` or both if you would like to have some additional experience running both tools. FYI, `DataExplorer` does create an html file when you run it and does not create an output within script.

#### Questions

##### Discuss the shape of the distribution for a few of your numeric features? For example, `dep_delay` appers to be positively skewed. 
**Answer:** 

```{r}
flights %>% skimr::skim()

# Set up for DataExplorer
#config <- configure_report(
#  add_plot_str = FALSE,
#  add_plot_qq = FALSE,
#  add_plot_prcomp = FALSE,
#  add_plot_boxplot = FALSE,
#  add_plot_scatterplot = FALSE,
#  global_ggtheme = quote(theme_minimal(base_size = 14))
#)
#create_report(flights, config = config)
```

##### The distribution plots are some times hard to read when they presented as a summary. Dig into two columns of interest and discuss what else you learned from the additional plot. Explore the charts and graphs discussed in chapter two of Sharda or chapter 7 of Lander. For example, my assertion that the delay variables for arrival time and departure time do appear to be positively skewed. 

**Answer:** 

```{r}
#Create 2 histograms showing arr_delay and dep_delay across 2013 flights
par(mfrow=c(1,2))
hist(flights$arr_delay, main = "Arrival Time Delays")
hist(flights$dep_delay, main = "Departure Time Delays")
```